---
title: "CSS 205/POLI 271 Final Project"
author: "Amy Tan"
format: pdf
editor: visual
---

## Load the data for your replication project into R. Produce the following:

```{r}
library(haven)
library(tidyverse)
study2_df <- read_sav('original_data/Study2.sav')

```

1.  A histogram of the dependent variable

    ```{r}
    library(ggplot2)

    ggplot(study2_df, aes(x=control)) + 
      geom_bar()

    ```

```{r}
ggplot(study2_df, aes(x=suma)) + 
  geom_bar()
```

```{r}
ggplot(study2_df, aes(x=praca)) + 
  geom_bar()
```

```{r}
ggplot(study2_df, aes(x=potreba)) + 
  geom_bar()
```

```{r}
ggplot(study2_df, aes(x=pila)) + 
  geom_bar()
```

2.  A correlation matrix for the DV and IVs that the original authors included in the model you are replicating

    ```{r}

    ```

3.  A visual or tabular depiction of the missingness in the data from part (2); see p. 251-255 of the text.

    ```{r}

    ```

Submit this output as a pdf file.

## Data

Accessed from [Dataverse](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/6KCW8Z).
