---
title: "CSS 205/POLI 271 Final Project"
author: "Amy Tan"
format: pdf
editor: visual
---

## Load the data for your replication project into R. Produce the following:

```{r}
# importing relevant libraries
library(haven)
library(tidyverse)
library(ggplot2)

# Saving study 2 data as s2
s2 <- read_sav('original_data/Study2.sav')

# Data cleaning code provided from authors' Dataverse Rmd

# Convert the 'home' column into a factor.
# Levels 1-4 are mapped to descriptive labels for different types of responses.
s2$home= factor(s2$home, 
                levels = c(1, 2,3,4), 
                labels = c("home-answered","away-notanswered","home-return", "away-return"))

# Convert the 'consent' column into a factor.
# Level 1 becomes "yes" and level 2 becomes "no".
s2$consent= factor(s2$consent, 
                   levels = c(1, 2), 
                   labels = c("yes","no"))

# Convert the 'etnic_ascr' column into a factor.
# Level 1 is labeled "slovak" and level 2 is labeled "roma".
s2$etnic_ascr= factor(s2$etnic_ascr, 
                      levels = c(1, 2), 
                      labels = c("slovak","roma"))

# Convert the 'etnicita' column into a factor.
# Levels 1, 2, and 3 are mapped to "slovak", "roma", and "other" respectively.
s2$etnicita= factor(s2$etnicita, 
                    levels = c(1, 2,3), 
                    labels = c("slovak","roma","other"))

# Ensure that the 'age' column is numeric.
s2$age <- as.numeric(s2$age)

# Create a lookup vector mapping ages to years.
# For each value in 'age' from 9 to 72, recode it to a corresponding year from 1996 to 1933.
# The result is stored in a new variable called 'roky'.
lookup <- setNames(1996:1933, 9:72)
s2$roky <- lookup[as.character(s2$age)]

# Calculate a new variable 'years' representing the difference between 2019 and the recoded year.
s2$years <- 2019 - s2$roky

# Convert the 'gender' column into a factor.
# Level 1 is labeled "Male" and level 2 is labeled "Female".
s2$gender <- factor(s2$gender, 
                    levels = c(1, 2), 
                    labels = c("Male","Female"))

# Create a subset of the data containing only rows where consent is "yes".
s2_consent <- dplyr::filter(s2, consent=="yes")

# Further filter the subset to exclude rows where 'etnic_ascr' is "NA".
s2_consent <- dplyr::filter(s2_consent, etnic_ascr != "NA")

# Further filter the subset to only include rows where multiple survey items (e.g., eurofondy_iv, skolka_iv, etc.) have values less than 5. 
s2_consent <- dplyr::filter(s2_consent, eurofondy_iv<5 & skolka_iv<5 & skolka_agree<5 & skolka_vote<5 & skolka_norms < 5 & control < 5 & suma < 5 & praca < 5 & potreba<5 & pila <5 )
```

1.  A histogram of the dependent variable

    ```{r}


    ```

```{r}
ggplot(study2_df, aes(x=suma)) + 
  geom_bar()
```

```{r}
ggplot(study2_df, aes(x=praca)) + 
  geom_bar()
```

```{r}
ggplot(study2_df, aes(x=potreba)) + 
  geom_bar()
```

```{r}
ggplot(study2_df, aes(x=pila)) + 
  geom_bar()
```

2.  A correlation matrix for the DV and IVs that the original authors included in the model you are replicating

    ```{r}

    ```

3.  A visual or tabular depiction of the missingness in the data from part (2); see p. 251-255 of the text.

    ```{r}

    ```

Submit this output as a pdf file.

## Data

Accessed from [Dataverse](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/6KCW8Z).
